### 2014-09-11 paper/paper.Rmd
occupation

# 5. Parents matter!
### a. Number of parents in the home
Looking at the number of parents in a household shows that students in two-parent households are on average performing better than students not in a two-parent households. Students with just a mother at home do tend to perform better than students with only a father at home. Students with no parents at home are performing the worst, on average.

@@

### b. Parent's Occupation Status
One of the variables the survey addressed was the occupation status of each of the parents. There were five possible categories: “Full-Time”, “Part-Time”, “Not working, but looking for a job”, “Other (e.g. home duties, retired)”, and no response. The distribution of the number of parents who selected each of the categories can be seen in the bar chart. The chart shows that fathers primarily have full-time jobs. However, there are large numbers of mothers who work full-time, part-time, or are stay-at-home moms. With women making up a larger portion of the work-force, it is interesting to look at how the parents' occupations relate to students' performance in mathematics, reading, and science.

The plot below shows the average performance of students' performance based on the parents' occupational statuses. The countries are sorted according to the average score for students with a "stay-at-home" mother. The averages for each parent in each of the work status categories are distinguished by color.  In each of the plots below, students with mothers who work part-time or full-time tend outperform those students with stay-at-home mothers. This pattern holds with only a few exceptions. Meanwhile, when the father is employed part-time, students tend to have much lower achievement that if the father stayed at home. Those students whose fathers have full-time jobs have the highest average achievement among the possible work statuses for the father. The patterns described hold across mathematics, reading, and science.

@@

a8 <- rbind(cbind(a6,Parent="Mother"),cbind(a7,Parent="Father"))
#a2 <- melt(a1)
ind <- which(a8$Job.Status=="Full-Time"& a8$Parent=="Mother")
a8$name <- factor(a8$name,levels(a8$name)[order(a8$Math[ind])])
p <- qplot(name, Math, col=Job.Status, data = a8) +coord_flip() + facet_wrap(~Parent)
p
```

>>>>

# 5. Parents matter!
### a. Number of parents in the home
The role of parents in educational achievement is the subject of much research. Looking at the number of parents in a household shows that students in two-parent households are, on average, performing better in math than students not in a two-parent households. The majority of students live in two-parent households. Of the single-parent households, a single mother is more common than a single father. Scores of students with just a mother at home tend to be higher than those with just a father at home. In many countries, mostly Asian, students from single-mother households achieve comparably to those from two-parent households. Students in households with no parents tend to perform the worst. This pattern is consistent across all countries.

@@

### b. Parent's Occupation Status
One of the variables the survey addressed was the occupation status of each of the parents. There were five possible categories: “Full-Time”, “Part-Time”, “Not working, but looking for a job”, “Other (e.g. home duties, retired)”, and no response. The frequencies for these categories can be seen in the bar chart. The chart shows that fathers primarily have full-time jobs. However, there are large numbers of mothers who work full-time, part-time, or are stay-at-home moms. With women making up a larger portion of the work-force, it is interesting to look at how the parents' occupations relate to students' performance in math.

The plot below shows the average student math score by the parents' occupational status. The countries are sorted according to the average score for students with a "full-time" mother. Occupational status are distinguished by color.  In each of the plots below, students with mothers who work part-time or full-time tend outperform those students with stay-at-home mothers. This pattern holds with only a few exceptions. Meanwhile, when the father is employed part-time, students tend to have much lower achievement that if the father stayed at home. Those students whose fathers have full-time jobs have the highest average achievement among the possible work statuses for the father. The patterns described hold across mathematics, reading, and science.

@@

a8 <- rbind(cbind(a6,Parent="Mother"),cbind(a7,Parent="Father"))
#a2 <- melt(a1)
ind <- which(a8$Job.Status=="Full-Time"& a8$Parent=="Mother")
a8$name <- factor(a8$name,levels(a8$name)[order(a8$Math[ind])])
p <- qplot(name, Math, col=Job.Status, data = a8, xlab="") + coord_flip() + facet_wrap(~Parent) + theme(legend.position="bottom")
p
```
